On the error performance of the $A_n$ lattices

### 2012-01-23 paper.tex
minor fixes to paper.

The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}. The highly symmetric structure of these lattices provides the grounds for extremely efficient decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.  In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.  This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.  Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices,~\cite{Conway1982VoronoiRegions}.  In this paper we extend their technique to compute the moments of any order for $A_n$.

In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.  These are the densest known sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.  In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but does offer a coding gain in small dimensions.  For these cases, we provide an error probability expression that can be computed to any degree of accuracy at any finite signal-to-noise ratio.
%We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.  In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result (without proof), which gives recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $n+1$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
 %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $n$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$.  We also plot the results of Monte-Carlo simulations that support our analytical results.
%An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis. In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}. A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix. It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.

The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\mathcal{S}_{\Lambda}$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point. It can be shown that the Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.  It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.  That is, the Voronoi cell tessellates space when translated by points in $\Lambda$.  To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed. Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.  We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.

@@

Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.  Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).  A result of this is that that observed noise distribution is no longer Gaussian. However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.  In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
\begin{align}
P_C &\leq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
 \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
\vor(\Lambda) + \xbf )  \nonumber \\
&= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
\end{align}
The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
\begin{align}
P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
4\sigma^42!} - \dots d\xbf \nonumber \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
d\xbf.  \label{eq:summomentproberror}
\end{align}
So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.  The number of terms required will increase as the noise variance gets smaller.  This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.

@@

We now state our main result.  The $m$th moment $\calM_n(m)$ of the lattice $A_n$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{(-1)^{k-a}H(n,m,k,a,b)}
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1}
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
\begin{align*}
\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor{A_n}$~ \cite[p. 462]{SPLAG}}, \\
\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have explicitly tabulated these formula for $m=0$ to $40$. For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.  We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.

@@

The Voronoi cell of $A_n$ is closely related to the $n+1$ dimensional hypercube $\vor(\ints^{n+1})$ as the next theorem will show.  This result has appeared previously~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}, but we repeat it here so that this paper is self contained.  We denote by
\[
\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{\onebf^\prime \onebf} = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
\]
the projection matrix orthogonal to $\onebf$ (i.e. into the zero-sum plane) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.  Given a set $S$ of vectors from $\reals^{n+1}$ we write $\Qbf S$  to denote the set with elements $\Qbf \sbf$ for all $\sbf \in S$, i.e. the set containing the projection of the vectors from $S$.

@@

We would like to be able to integrate functions over the Voronoi cell of $A_n$.  Consider a function $f : \reals^{n+1} \mapsto \reals$.  The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.  The Voronoi cell is a subset of the zero-sum plane that has zero $n+1$-dimensional volume.  So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.  This is not what we intend.  By an appropriate change of variables it would be possible to write the Voronoi cell $\vor(A_n)$ in an $n$-dimensional coordinate system, and then integrate.  However, we find the following approach simpler.  Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.  Now, the integral over the Voronoi cell can be written as
\begin{equation}\label{eq:vorfint}
\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
\end{equation}
It is not immediately clear how an integral over $\vor(A_n) \times \onebf$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
\begin{equation}\label{eq:unormalised}
\int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
\end{equation}
This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.  To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.  Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.  The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by
\[
\|\onebf\|(1 - x_{\text{max}} + x_{\text{min}}) = \sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}}).
\]
The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
\[
\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
\]
The primary advantage of this integral is that the bounds are given by the hypercube $\vor(\ints^{n+1})$.

@@

We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables gives
\[
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} \frac{f(\Qbf\xbf )}{1 - t}  dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
\]
and letting $w_{i-2} = x_i - y + 1/2 + t$ for $i = 3,\dots,n+1$ gives
\begin{equation}\label{eq:intwewilluse}
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{t}_{0} \cdots \int^{t}_{0} \frac{f(\Qbf\xbf )}{1 - t}  dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
\end{equation}
Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector
\[
\wbf = [t, 0, w_1, w_2, \dots, w_{n-1}]^\prime.
\]
Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.  Interestingly $\wbf$ does not contain $y$ so the term inside the integral does not depend on $y$.  This is the integral we will use to compute the moments of $A_n$.

@@

The plot also display an approximation for the probability of error for the 8-dimensional $E_8$ lattice.  The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.  The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.  The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.  Applying the union bound the probability of error satisfies
\newcommand{\erfc}{\operatorname{erfc}}
\newcommand{\erf}{\operatorname{erf}}
\[
P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
\]
where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.  For the $E_8$ lattice this approximation is an upper bound because the relevant vectors of $E_8$ (those vectors that define the Voronoi cell) are precisely the 240 minimal vectors.  %The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.

>>>>

The \emph{root lattices} $A_n$, $D_n$, $E_6$, $E_7$ and $E_8$ have attracted particular attention as structured codes for the AWGN channel~\cite{SPLAG}. The highly symmetric structure of these lattices provides the grounds for extremely efficient encoding and decoding algorithms~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.  In this paper we consider codes constructed from the root lattice $A_n$ and derive formulae for accurately predicting the performance of these codes.  This is achieved by deriving formulae for the \emph{moments} of the \emph{Voronoi cell} of $A_n$.  Conway and Sloane suggested this approach to compute the quantizing constants (second order moments) of the root lattices~\cite{Conway1982VoronoiRegions}.  In this paper we extend their technique to compute the moments of any order for $A_n$.

In two dimensions $A_2$ is the hexagonal lattice and in three dimensions $A_3$ is the face-centered cubic lattice.  These are the densest sphere packings in dimensions two and three and our results automatically include low dimensional codes constructed using these packings.  In general, the lattice $A_n$ does not produce asymptotically good codes in large dimensions, but does offer a coding gain in small dimensions.  For these cases, we provide an error probability expression that can be computed to any degree of accuracy at any finite signal-to-noise ratio.
%We are hopeful that the techniques developed here may be applied to other lattices that do produce good codes.  In particular our techniques might apply to the Coxeter lattices $A_n^m$ of which the root lattice $E_8 \simeq A_8^3$ is a member~\cite{Coxeter1951,Martinet2003,McKilliam2009CoxeterLattices}.

This paper is organised as follows.  In Section~\ref{sec:latt-latt-codes} we give a brief overview of lattices and codes constructed from them, i.e., \emph{lattice codes}.  We describe lattice decoding and show how the probability of coding error can be expressed in terms of the moments of the \emph{Voronoi cell} of the lattice used.  Section~\ref{sec:main-result} states the main result, describing recursive formula to compute the moments of the Voronoi cell of $A_n$.  Section~\ref{sec:lattice-a_n} describes the lattice $A_n$ and some of its properties.  An important property for our purposes is that the Voronoi cell of $A_n$ is precisely the projection of a $(n+1)$-dimensional hypercube orthogonal to one of its vertices~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.
 %This result was previously known~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}.  The proof is short and we give it again here so that this paper is self contained.
In Section~\ref{sec:integr-funct-over} we use this property to show how integrals over the Voronoi cell of $A_n$ can be expressed as integrals over the $(n+1)$-dimensional hypercube. These integrals are solvable and we use them to obtain the moments of the Voronoi cell in Section~\ref{sec:powers-eucl-norm}.  In Section~\ref{sec:results-simulations} we plot the probability of error versus signal to noise ratio for codes constructed from the lattices $A_1 \simeq \ints$, $A_2$, $A_3 \simeq D_3$ and $A_4$, $A_5$ and $A_8$.  We also plot the results of Monte-Carlo simulations that support our analytical results.
%An advantage of the analytical results is that the probability of error can calculated at high SNR, when the probabilty of error is small.  This is in contrast to Monte-Carlo simulations that are computationally infeasible at when the probability of error is small.  We display the performance of these lattice codes

@@

A \term{lattice}\index{lattice}, $\Lambda$, is a discrete subset of points in $\reals^m$ such that
\[
   \Lambda = \{\xbf = \Bbf\ubf \mid \ubf \in \ints^n \}
 \]
 where $\Bbf \in \reals^{m \times n}$ is an $m \times n$ matrix of rank $n$, called the \term{generator matrix} or \term{basis matrix} or simply generator or basis. In particular, the set of $n$-tuples of integers $\ints^n$ is a lattice (with the identity matrix as the generator) and we call this the~\emph{integer lattice}. A lattice $\Lambda$ associated with a rank-$n$ generator matrix $\mathbf{B}$ is said to be $n$-dimensional. If the generator is square, i.e. $m = n$, then the lattice points span $\reals^n$ and we say that the lattice is \term{full rank}. If $\Bbf$ has more rows than columns, i.e. $m > n$, then the lattice points lie in a $n$-dimensional subspace of $\reals^m$. For any lattice $\Lambda$ with an $m \times n$ generator matrix, we define $\mathcal{S}_{\Lambda}$ to be the hyperplane spanned by the columns of the generator matrix. %It is easy to see that $\mathcal{S}_\Lambda$ is then the $n$-dimensional subspace containing the lattice $\Lambda$.

The (open) \term{Voronoi cell}, denoted $\vor(\Lambda)$, of a lattice $\Lambda$ is the subset of $\mathcal{S}_{\Lambda}$ containing all points nearer (in Euclidean distance) to the lattice point at the origin than any other lattice point. The Voronoi cell is an $n$-dimensional convex polytope that is symmetric about the origin.  It is convenient to modify this definition of the Voronoi cell slightly so that the union of translated Voronoi cells $\cup_{\xbf \in \Lambda}\vor(\Lambda) + \xbf$ is equal to  $\mathcal{S}_{\Lambda}$.  That is, the Voronoi cell tessellates when translated by points in $\Lambda$.  To ensure this we require that if a face of $\vor(\Lambda)$ is open, then its opposing face is closed. Specifically, if $\xbf \in \vor(\Lambda)$ is on the boundary of $\vor(\Lambda)$ then $-\xbf \notin \vor(\Lambda)$.  We wont specifically define which opposing face is open and which is closed as the results that follow hold for any choice of open and closed opposing faces.

@@

Let $\calX$ be a lattice code constructed from a lattice $\Lambda$.  Assuming that each codeword $\xbf \in \calX$ has an equal probability of being transmitted and that the receiver employs lattice decoding, then the average probability of correct decoding in the AWGN channel is
%\footnote{In order to approach capacity Erez and Zamir~\cite{Erex2004_lattice_decoding} employ a uniform dither and also a minimum mean square error scaling parameter $\alpha$ (the so called \emph{modulo-lattice additive noise channel}).  A result of this is that that observed noise distribution is no longer Gaussian. However, for large signal to noise ratio $\alpha \approx 1$ and the observed noise is very close to Gaussian.  In this case~\eqref{eq:PEgaussian} is a reasonable approximation.}
\begin{align}
P_C &\geq P_C(\Lambda) = \lim_{\abs{\calX}\rightarrow \infty}
 \frac{1}{\abs{\calX}} \sum_{\xbf\in \calX} \prob( \xbf + \wbf \in
\vor(\Lambda) + \xbf )  \nonumber \\
&= \prob(\xbf \in \vor(\Lambda) ) \nonumber \\
&=   \frac{1}{(\sqrt{2\pi}\sigma)^n} \int_{\vor(\Lambda)}
e^{-\|\xbf\|^2 / 2\sigma^2 } d\xbf. \label{eq:PEgaussian}
\end{align}
The probability of error is $P_E \leq 1 - P_C(\Lambda)$, where the upper bound is asymptotically tight for large constellations (the proportion of the codewords near the boundary of the shaping region becomes small).  By expanding $e^x = 1  + x + \frac{x^2}{2} + \dots$ according to its Maclaurin series we obtain
\begin{align}
P_C(\Lambda)  &= \frac{1}{(\sqrt{2\pi}\sigma)^n}\int_{\vor(\Lambda)} 1
- \frac{\|\xbf\|^2}{2\sigma^2} + \frac{\left(\|\xbf\|^2\right)^2}{
4\sigma^42!} - \dots d\xbf \nonumber \\
&= \frac{1}{(\sqrt{2\pi}\sigma)^n} \sum_{m=0}^\infty
\frac{(-1)^m}{2^m\sigma^{2m}m!} \int_{\vor(\Lambda)} \|\xbf\|^{2m}
d\xbf.  \label{eq:summomentproberror}
\end{align}
So, to obtain arbitrarily accurate approximations to the probability of error it is enough to know the values of $\int_{\vor(\Lambda)} \|\xbf\|^{2m} d\xbf$ for $m=1,2\dots$ for some sufficiently large $m$.  The number of terms required will increase as the noise variance gets smaller.  This implies that the bound with a fixed number of terms is not asymptotically tight but is very accurate up to a finite signal-to-noise ratio.

@@

We now state our main result.  The $m$th moment $\calM_n(m)$ of the lattice $A_n$ satisfies
\begin{equation}\label{eq:theCmformula}
\frac{\calM_n(m)}{m!} = \frac{n\sqrt{n+1}}{n+2m}\sum_{k=0}^{m}\sum_{a =0}^{k}\sum_{b=0}^{k-a} \frac{G(n-1,a,2k - 2a - b)}{H(n,m,k,a,b)},
\end{equation}
where the function
\[
H(n,m,k,a,b) = \frac{(n+1)^{m-a}a!(m-k)!b! (k-a-b)!}{(-1)^{k-a}2^{b} n^{m-k}},
\]
and the function $G(n,c,d)$ satisfies the recursion
\begin{equation}\label{eq:theGrecursion}
G(n,c,d) = \sum_{c'=0}^{c} \sum_{d'=0}^{d} \binom{c}{c'}\binom{d}{d'} \frac{G(n-1,c-c',d-d')}{2c'+d'+1},
\end{equation}
with the initial conditions
\[
G(1,c,d) = \frac{1}{2c+d+1} \qquad \text{and} \qquad G(n,0,0) = 1.
\]
For fixed $m$ it is possible to solve this recursion in $n$ and obtain formula for the $\calM_n(m)$ in terms of $n$.  The first five such formula are:
\begin{align*}
\calM_n(0) &= \sqrt{n+1} \qquad \text{the volume of } \vor(A_n),\\
\calM_n(1) &= \frac{n(n+3)}{12\sqrt{n+1}} \qquad \text{the second moment of $\vor(A_n)$~ \cite[p. 462]{SPLAG}}, \\
\calM_n(2) &=  \frac{50 n + 55 n^2 + 34 n^3 + 5 n^4}{720 (1 + n)^{3/2}}, \\
\calM_n(3) &= \frac{1960 n + 2142 n^2 + 2681 n^3 + 1423 n^4 + 399 n^5 + 35 n^6}{60480 (1 + n)^{5/2}}, \\
\calM_n(4) &= \frac{93744 n + 34356 n^2 + 112172 n^3 + 89343 n^4 + 53224 n^5 + 17246 n^6 + 2940 n^7 + 175 n^8}{3628800 (1 + n)^{7/2}}.
\end{align*}
We have explicitly tabulated these formula for $m=0$ to $40$. For larger $m$ direct evaluation for specific $n$ from the recursive formula is preferable.  We will derive these results in Section~\ref{sec:powers-eucl-norm}, but first need some properties of the lattice $A_n$.

@@

The Voronoi cell of $A_n$ is closely related to the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$ as the next theorem will show.  This result has appeared previously~\cite{McKilliam2009CoxeterLattices,McKilliam2010thesis}, but we repeat it here so that this paper is self contained.  We denote by
\[
\Qbf = \Ibf - \frac{\onebf\onebf^\prime}{\onebf^\prime \onebf} = \Ibf - \frac{\onebf\onebf^\prime}{n+1}
\]
the projection matrix orthogonal to $\onebf$ (i.e. into the zero-sum plane) where $\Ibf$ is the $n+1$ by $n+1$ identity matrix.  Given a set $S$ of vectors from $\reals^{n+1}$ we write $\Qbf S$  to denote the set with elements $\Qbf \sbf$ for all $\sbf \in S$, i.e. the set containing the projection of the vectors from $S$.

@@

We would like to be able to integrate functions over the Voronoi cell of $A_n$.  Consider a function $f : \reals^{n+1} \mapsto \reals$.  The definition we have made for $A_n$ above places it in the $n$-dimensional zero-sum plane, lying in $\reals^{n+1}$.  The Voronoi cell is a subset of the zero-sum plane that has zero $(n+1)$-dimensional volume.  So, the integral $\int_{\vor(A_n)} f(\xbf) d \xbf$ is equal to zero.  This is not what we intend.  By an appropriate change of variables it would be possible to write the Voronoi cell $\vor(A_n)$ in an $n$-dimensional coordinate system, and then integrate.  However, we find the following approach simpler.  Given a set $S$ of vectors from the zero sum plane, let $S \times \onebf$ denote the set of elements that can be written as $\xbf + \ybf$ where $\xbf \in S$ and $\ybf = k\onebf$ for some $k \in [-\nicefrac{1}{2},\nicefrac{1}{2}]$.  Now, the integral over the Voronoi cell can be written as
\begin{equation}\label{eq:vorfint}
\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf = \int_{\Qbf\vor(\ints^{n+1}) \times \onebf} f(\Qbf\xbf) d\xbf.
\end{equation}
It is not immediately clear how an integral over $\vor(A_n) \times \onebf$ should be performed.  Consider the following simpler integral over the hypercube $\vor(\ints^{n+1})$,
\begin{equation}\label{eq:unormalised}
\int_{\vor(\ints^{n+1})} f(\Qbf\xbf) d\xbf.
\end{equation}
This integral is not equal to~\eqref{eq:vorfint} because, although $\Qbf\xbf$ is always an element of $\vor(A_n)$, the integral is not uniform over $\vor(A_n)$.  To see this, consider some $\xbf \in \vor(\ints^{n+1})$ and let $x_{\text{max}}$ be the maximum element of $\xbf$ and $x_{\text{min}}$ be the minimum element.  Then $\xbf +  k \onebf \in \vor(\ints^{n+1})$ for those $k \in [-\nicefrac{1}{2} - x_{\text{min}}, \nicefrac{1}{2} - x_{\text{max}})$.  The length of this interval is $1 - x_{\text{max}} + x_{\text{min}}$ so the (one dimensional) volume of points in $\vor(\ints^{n+1})$ that, once projected orthogonally to $\onebf$, are equal to $\Qbf\xbf$ is given by
\[
\|\onebf\|(1 - x_{\text{max}} + x_{\text{min}}) = \sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}}).
\]
The integral~\eqref{eq:vorfint} can be obtained by normalising~\eqref{eq:unormalised} by this length, that is,
\[
\int_{\vor(A_n) \times \onebf} f(\Qbf\xbf) d\xbf =  \int_{\vor(\ints^{n+1})} \frac{f(\Qbf\xbf)}{\sqrt{n+1}(1 - x_{\text{max}} + x_{\text{min}})}  d\xbf.
\]
The primary advantage of this integral is that the bounds are given by the $(n+1)$-dimensional hypercube $\vor(\ints^{n+1})$.

@@

We can make further simplifications.  Letting $t = x_1 - x_2$ and $y = x_1 + 1/2$ and changing variables,
\[
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{y - 1/2}_{y - t - 1/2} \cdots \int^{y-1/2}_{y-t-1/2} \frac{f(\Qbf\xbf )}{1 - t}  dx_{n+1} \, \dots \, dx_3 \, dt \, dy,
\]
and letting $w_{i-2} = x_i - y + 1/2 + t$ for $i = 3,\dots,n+1$ gives
\begin{equation}\label{eq:intwewilluse}
n \sqrt{n+1} \int^{1}_{0} \int^{y}_{0}  \int^{t}_{0} \cdots \int^{t}_{0} \frac{f(\Qbf\xbf )}{1 - t}  dw_{n-1} \, \dots \, dw_1 \, dt \, dy.
\end{equation}
Observe that $\xbf = \wbf + (y - t - \nicefrac{1}{2})\onebf$ where $\wbf$ is the column vector
\[
\wbf = [t, 0, w_1, w_2, \dots, w_{n-1}]^\prime.
\]
Projecting orthogonal to $\onebf$ gives $\Qbf\xbf = \Qbf \wbf$.  Interestingly $\wbf$ does not contain $y$ so the term inside the integral~\eqref{eq:intwewilluse} does not depend on $y$.  This is the integral we will use to compute the moments of $A_n$.

@@

The plot also displays an approximation for the probability of error for the 8-dimensional $E_8$ lattice.  The approximation is made in the usual way by applying the union bound to the minimal vectors of the lattice~\cite[p.~71]{SPLAG}.  The $E_8$ lattice has 240 minimal vectors of length $\sqrt{2}$.  The packing radius of $E_8$ is therefore $\rho = \sqrt{2}/2$.  Applying the union bound the probability of error satisfies
\newcommand{\erfc}{\operatorname{erfc}}
\newcommand{\erf}{\operatorname{erf}}
\[
P_E \leq 240\erfc\left( \frac{\rho}{\sqrt{2}\sigma} \right) = 240\erfc\left(\frac{1}{2\sigma}\right)
\]
where $\erfc(x) = 1 - \erf(x)$ is the complementary error function.  For the $E_8$ lattice this approximation is an upper bound because the relevant vectors of $E_8$ (those vectors that define the Voronoi cell) are precisely the 240 minimal vectors.  %The second moment of this representation of $E_8$ is $\frac{929}{1620}$ so SNR is related to $\sigma^2$ by $\text{SNR} = \frac{929}{3240\sigma^2}$.
